Figures for 3-D printed oxygen devices

### 2015-08-13 publishing/response-to-reviewers.md

PONE-D-15-17915

3D-Printed 24-Well Plate Oxygen Control Device

PLOS ONE

@@

In addition to this rebuttal letter a clean revised manuscript is submitted as well as a marked-up copy with tracked changes.

Yours sincerely,

David Eddington, Ph.D.
University of Illinois at Chicago

**Authors' Rebuttal to Reviewers:**

**Reviewer #1:**

It is a different version of design to control oxygen in cell culture, among some recent efforts. One concern is that the group has published quite a number of papers on oxygen control devices, although it is now through 3d printing. It is important to highlight in front the key improvements.

* **Authors' rebuttal:** 
An overview of improvements from our previous PDMS cast devices starts on line 51 of the original manuscript, now line 64 of the revised manuscript.

@@

Also, it is important to compare with some other methods related to 3d printing such as that done recently by Kamei K. et al. (Biomed Microdevices. 2015 Apr;17(2):36).

* **Authors' rebuttal:** 
The work by Kamei et al. was cited along with several other papers that used 3D printed parts as molds for PDMS devices on line 33 of the original manuscript, now line 46 of the revised manuscipt.

@@

"This is the first 3D-printed device that can be functionalized to control oxygen in cell culture" is a statement which is too strong.

* **Authors' rebuttal:** 
The wording was changed to "This is first 3D-printed device incorporating gas permeable membranes to facilitate oxygen control in cell culture."

**Reviewer #2:**

Currently, little consideration is given to mimicking physiological oxygen concentrations in vitro during cell experiments. Oftentimes cells are cultured at atmospheric oxygen conditions which are higher than what is found in many tissues, including tumor environments. Some systems have been developed to control the oxygen environment in vitro, but these can be cumbersome or costly to implement. 3D printing has emerged as a rapid prototyping method to fabricate devices with high control over structure and therefore is an attractive method to design a microfluidic device capable of controlling oxygen concentrations in culture. Brennan et al. report on a 3D printed control device capable of adjusting oxygen concentrations in a 24-well plate. Although the concept is interesting, the data is not sufficient and additional experiments are needed to improve the manuscript.

Page 3 - In addition to presenting gap distance below insert and well bottom, it would be helpful if the authors included volume of media this space would hold.

* **Authors' rebuttal:**
A sentence was added including the approximate volume of media in the gap. See line 87 of the revised manuscript.

Page 4 – Why was 6 hrs chosen for cell culture? It would seem that the authors could demonstrate a more robust change in expression over the concentrations tested if the experiment was extended to a longer time point, especially considering only the 0% group showed a significant difference in VEGFA expression.

* **Authors' rebuttal:** 
VEGFA is under the transcriptional activity of HIF-1alpha. HIF protein peaks at 4 hr, and so we looked at a time point slightly delayed. By longer time points, the HIF response is actually suppressed, and many transcripts are not increased with longer time points. See Rexius-Hall et al.s' work, citation 44.

The authors specify concentrations of 21%, 10%, 5%, and 0% were evaluated, however in their Figure 2b, the concentrations appear closer to 13% (not 10%) and 7% (not 5%). Variation of 3% can be considerable in physiological environments and it is unclear if the authors recognize this variation. It is also amplified by the lack of error bars for each group.

* **Authors' rebuttal:**
The data in the original submission was analyzed with a two point calibration. This method correlates intensities to oxygen concentration assuming a linear Stern-Volmer relationship. While standard practice, the two-point calibration suffers from increasing inaccuracy as points are measured further away from the calibration points. Characterization data was re-taken with 5 calibration points and analyzed with a two-site Stern-Volmer model to increase accuracy of analyzed data. A new plot replaces figure 2 and includes error bars.

Is sterilization an issue for this device, considering it will be in close contact with biological components?

* **Authors' rebuttal:**
Sterilization was preformed by spraying the device with 70% ethonal and leaving it in a biosaftey hood with UV lights overnight. We did not observe any contamination. This information was added to the manuscript. See lines 131, and 141.

### 2015-08-17 publishing/response-to-reviewers.md

It is a different version of design to control oxygen in cell culture, among some recent efforts. One concern is that the group has published quite a number of papers on oxygen control devices, although it is now through 3d printing. It is important to highlight in front the key improvements.

* **Authors' rebuttal:** 
An overview of improvements from our previous PDMS cast devices starts on line 51 of the original manuscript, now line 64 of the revised manuscript.

Also, it is important to compare with some other methods related to 3d printing such as that done recently by Kamei K. et al. (Biomed Microdevices. 2015 Apr;17(2):36).

* **Authors' rebuttal:** 
The work by Kamei et al. was cited along with several other papers that used 3D printed parts as molds for PDMS devices on line 33 of the original manuscript, now line 46 of the revised manuscript.

@@

Page 3 - In addition to presenting gap distance below insert and well bottom, it would be helpful if the authors included volume of media this space would hold.

* **Authors' rebuttal:**
A sentence was added including the approximate volume of media in the gap. See line 87 of the revised manuscript.

Page 4 – Why was 6 hrs chosen for cell culture? It would seem that the authors could demonstrate a more robust change in expression over the concentrations tested if the experiment was extended to a longer time point, especially considering only the 0% group showed a significant difference in VEGFA expression.

* **Authors' rebuttal:** 
VEGFA is under the transcriptional activity of HIF-1alpha. The time course of HIF-1alpha protein levels in A549 cells demonstrates maximal protein levels at 4 h and a steep decrease in expression by 16 h. Based on the known time course, we looked at a time point slightly delayed from the 4 h peak to account for transcription of the downstream target gene of interest (VEGFA). At longer time points, HIF protein expression is actually suppressed, and many transcripts are not increased with extended duration. This additional information was added along with a citation starting at line 182.

@@

The authors specify concentrations of 21%, 10%, 5%, and 0% were evaluated, however in their Figure 2b, the concentrations appear closer to 13% (not 10%) and 7% (not 5%). Variation of 3% can be considerable in physiological environments and it is unclear if the authors recognize this variation. It is also amplified by the lack of error bars for each group.

* **Authors' rebuttal:**
The data in the original submission was analyzed with a two point calibration. This method correlates intensities to oxygen concentration assuming a linear Stern-Volmer relationship. While standard practice, the two-point calibration suffers from increasing inaccuracy as points are measured further away from the calibration points. Characterization data was re-taken with 5 calibration points and analyzed with a two-site Stern-Volmer model to increase accuracy of analyzed data. A new plot replaces figure 2 and includes error bars.

Is sterilization an issue for this device, considering it will be in close contact with biological components?

* **Authors' rebuttal:**
Sterilization was preformed by spraying the device with 70% ethanol and leaving it in a biosaftey hood with UV lights overnight. We did not observe any contamination. This information was added to the manuscript. See lines 131, and 141.

### 2015-08-17 publishing/latex/revision/revision.tex
added vegf data

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
\begin{flushleft}
{\Large
\textbf\newline{3D-Printed 24-Well Plate Oxygen Control Device}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Martin D. Brennan\textsuperscript{1},
Megan L. Rexius-Hall\textsuperscript{1},
David T. Eddington\textsuperscript{1,*},
\\
\bigskip
\bf{1} Dept of Bioengineering, University of Illinois at Chicago, Chicago, Illinois, USA
\\
\bigskip

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
3D printing has emerged as a method for directly printing complete microfluidic devices, although printing materials have been limited to oxygen-impermeable materials.
We demonstrate the addition of gas permeable PDMS (Polydimethylsiloxane) membranes to 3D-printed microfluidic devices as a means to enable oxygen control cell culture studies.
The incorporation of a 3D-printed device and gas-permeable membranes was demonstrated on a 24-well oxygen control device for standard multiwell plates.
The direct printing allows integrated distribution channels and device geometries not possible with traditional planar lithography.
With this device, four different oxygen conditions were able to be controlled, and six wells were maintained under each oxygen condition. 
We demonstrate enhanced transcription of the gene VEGFA (vascular endothelial growth factor A) with decreasing oxygen levels in human lung adenocarcinoma cells.
This is first 3D-printed device incorporating gas permeable membranes to facilitate oxygen control in cell culture.
\linenumbers

@@

Printing is currently limited in choice of substrate compatible with the 3D printing process.
Substrate options include many proprietary formulations which have been successfully used in a variety of applications.
New techniques for using 3D-printed molds to produce devices\cite{Mcdonald2002,Gregory2013,King2014,Comina2014a,Kamei2015} are also being developed, including fugitive ink methods\cite{Therriault2003,Miller2012,Bhargava2015}.
To date, there are no widely available methods or materials to facilitate direct printing of gas-permeable materials, although this area is actively being explored\cite{Femmer2014a}.
Microfluidic cell culture devices are most commonly cast in PDMS as it is a convenient material for cell studies due to its biocompatibility, optical properties, and gas permeability, facilitating oxygen control of cell environments\cite{Sia2003,Brennan2014}.

@@

Previously, we developed multiwell inserts for 6-well plates that controlled oxygen in a standard off-the shelf well plate \cite{Oppegard2009, Oppegard2010}.
These devices were completely cast from PDMS and fabricated in a multiple bonding procedure.
In addition, tubing was used to connect each pillar which became quite cumbersome when moving to a newer 24-well version.
Utilizing 3D printing for the fabrication allowed further features to be incorporated such as integrated distribution channels to eliminate the connection tubings between wells and hose barbs for better tubing connections, while eliminating fabrication time and failure rate.
Although available 3D printable materials are gas-impermeable, the addition of a PDMS membrane following printing of the passive microfluidic network is a simple addition allowing gas transfer. 
In this case, having a gas-impermeable material for the bulk is advantageous as it reduces unwanted gas transfer from the bulk material prior to the gas exchange interface.
The previous PDMS devices were Parylene coated along the convection channels to eliminate exchange of gas from the PDMS bulk, which again added to their complexity and is no longer needed.
Convection and distribution of the gas is done in the impermeable material, eliminating dilution of the gas before reaching the diffusion layer.

@@

Oxygen levels of 21\%, 10\%, 5\%, and 0\% were applied to each set of 6 wells in which A549 cells were cultured.
Measurement of relative gene expression confirmed biological responsiveness of cells under the device’s control.
The device’s ability to control the gas environment was demonstrated with hypoxia-induced upregulation of VEGFA mRNA in the A549 cells (Fig.~\ref{fig3}).  Quantitative PCR showed a significant increase in the relative gene expression of VEGFA in wells exposed to 0\% \ce{O2} for 6 hours as compared to 21\% \ce{O2}. Hypoxia-induced alterations in gene expression usually indicate a HIF-transduced hypoxic signal in cells.
When a HIF transcription factor undergoes nuclear translocation, it can initiate transcription of genes by binding to the hypoxia response element (HRE)\cite{Lee2004} on its target genes.
VEGFA has previously been shown to have a functional HRE, and upregulation has been observed in A549 cells under hypoxic conditions\cite{RohrerBley2011}.

\begin{figure}[h]
\caption{
{\bf PCR Data.}  VEGFA expression in A549 cells after exposure to different oxygen conditions set-up by the insert device. Data are expressed as mean $\pm$ SD, N=3.
}
\label{fig3}
\end{figure}

@@

\section*{Acknowledgments}
%This work funded by NSF 1253060.

>>>>

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
\begin{flushleft}
{\Large
\textbf\newline{A 3D-Printed Oxygen Control Insert for a 24-Well Plate}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Martin D. Brennan\textsuperscript{1},
Megan L. Rexius-Hall\textsuperscript{1},
David T. Eddington\textsuperscript{1,*},
\\
\bigskip
\bf{1} Dept of Bioengineering, University of Illinois at Chicago, Chicago, Illinois, USA
\\
\bigskip

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
3D printing has emerged as a method for directly printing complete microfluidic devices, although printing materials have been limited to oxygen-impermeable materials.
We demonstrate the addition of gas permeable PDMS (Polydimethylsiloxane) membranes to 3D-printed microfluidic devices as a means to enable oxygen control cell culture studies.
The incorporation of a 3D-printed device and gas-permeable membranes was demonstrated on a 24-well oxygen control device for standard multiwell plates.
The direct printing allows integrated distribution channels and device geometries not possible with traditional planar lithography.
With this device, four different oxygen conditions were able to be controlled, and six wells were maintained under each oxygen condition. 
We demonstrate enhanced transcription of the gene VEGFA (vascular endothelial growth factor A) with decreasing oxygen levels in human lung adenocarcinoma cells.
This is the first 3D-printed device incorporating gas permeable membranes to facilitate oxygen control in cell culture.
\linenumbers

@@

Printing is currently limited in choice of substrate compatible with the 3D printing process.
Substrate options include many proprietary formulations which have been successfully used in a variety of applications.
New techniques for using 3D-printed molds to produce devices\cite{Mcdonald2002,Gregory2013,King2014,Comina2014a,Kamei2015} are also being developed, including fugitive ink methods\cite{Therriault2003,Miller2012,Bhargava2015}.
To date, there are no widely available methods or materials to facilitate direct printing of gas-permeable materials, although this area is actively being explored\cite{Femmer2014a}.
Microfluidic cell culture devices are most commonly cast in PDMS as it is a convenient material for cell studies due to its biocompatibility, optical properties, and gas permeability, facilitating oxygen control of cell environments\cite{Sia2003,Brennan2014}.
In this study, a larger 24 well version was developed and optimized which includes several key improvements over the previous 6-well version.

@@

Previously, we developed multiwell inserts for 6-well plates that controlled oxygen in a standard off-the shelf well plate \cite{Oppegard2009, Oppegard2010}.
These 6-well devices were completely cast from PDMS and fabricated in a multiple bonding procedure.
In addition, tubing was used to connect each pillar which became quite cumbersome when moving to a newer 24-well version.
Utilizing 3D printing for the fabrication allowed further features to be incorporated such as integrated distribution channels to eliminate the connection tubings between wells and hose barbs for better tubing connections, while eliminating fabrication time and failure rate.
Although available 3D printable materials are gas-impermeable, the addition of a PDMS membrane following printing of the passive microfluidic network is a simple addition allowing gas transfer. 
In this case, having a gas-impermeable material for the bulk is advantageous as it reduces unwanted gas transfer from the bulk material prior to the gas exchange interface.
The previous PDMS devices were Parylene coated along the convection channels to eliminate exchange of gas from the PDMS bulk, which again added to their complexity and is no longer needed.
Convection and distribution of the gas is done in the impermeable material, eliminating dilution of the gas before reaching the diffusion layer.

@@

Oxygen levels of 21\%, 10\%, 5\%, and 0\% were applied to each set of 6 wells in which A549 cells were cultured.
Measurement of relative gene expression confirmed biological responsiveness of cells under the device’s control.
The device’s ability to control the gas environment was demonstrated with hypoxia-induced upregulation of VEGFA mRNA in the A549 cells (Fig.~\ref{fig3}).  Quantitative PCR showed a significant increase in the relative gene expression of VEGFA in wells exposed to 0\% \ce{O2} for 6 hours as compared to 21\% \ce{O2}. Hypoxia-induced alterations in gene expression usually indicate a HIF-transduced hypoxic signal in cells.
When a HIF transcription factor undergoes nuclear translocation, it can initiate transcription of genes by binding to the hypoxia response element (HRE)\cite{Lee2004} on its target genes.
VEGFA has previously been shown to have a functional HRE, and upregulation has been observed in A549 cells under hypoxic conditions\cite{RohrerBley2011}.
The 6 h time point was chosen based on the HIF-1$\alpha$ timecourse in A549 cells. 
HIF-1$\alpha$ protein levels are maximal at 4 h and steeply decrease by 16 h\cite{Li2006}. 
Based on the known time course, we looked at a time point slightly delayed from the 4 h peak to account for transcription of the downstream target genes.
\begin{figure}[h]
\caption{
{\bf PCR Data.}  VEGFA expression in A549 cells after exposure to different oxygen conditions set-up by the insert device. Data are expressed as mean $\pm$ SD, N=3.
}
\label{fig3}
\end{figure}

@@

%\section*{Acknowledgments}
%This work funded by NSF 1253060.
